Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-03-22 03_marlin_paper.qmd
tuna time

Many models have been developed to support the theory and design of spatial management strategies (see @fulton2015 for a review of existing methods). However, they have tended to fall into one of either highly complex and tactical models designed for use in a specific location, or to be highly stylized representations intended to provide theoretical insights with less relevance for specific applications. This lack of accessible models suited for more tactical application presents a challenge to stakeholders charged with marine spatial planning exercises. It also presents a barrier to the scientific community, wherein comparisons across spatial management simulation studies are clouded by discrepancies in underlying model structures beyond differences in the core phenomenon in question.

To help address this challenge, we present a novel bio-economic modeling tool called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient simulation of spatio-temporal biological and economic dynamics. `marlin` allows users to simulate the impacts of marine management policies across a range of species targeted by various fleets across heterogeneous and dynamic seascapes. This model can help users seek Parteo-optimal solutions that produce win-wins or minimize tradeoffs across multiple management objectives, such as food production and conservation [@lester2013].

@@

These species can be caught by fishing fleets. A fishing fleet is defined in `marlin` by a set of fishers that can have different fishing skill, prices, and selectivity for individual species (each of which we denote as a *metier*). For example, both a longline and purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can be made much more likely to catch bigeye than skipjack, and *vice versa*. Each fishing fleet distributes its fishing effort in space according to a specified spatial allocation rule, for example in proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.

Unconstrained by management, the total amount of effort exerted by each fleet can follow one of two dynamics: open access or constant effort. Under constant effort, the total amount of effort of each fleet is fixed over time, with the possible exception of attrition due to MPA placement. Under open access, the total amount of effort in a given time step is a function of the profitability of the fleet in the previous time step, evolving until a bionomic equilibrium of zero total profits is reached. Profitability is a function of the volume and price of each species caught, as well as the cost of the total amount of fishing effort per fleet and the travel costs as a function of distance from a port.

@@

For both the diffusion and taxis matrix, we allow for the inclusion of physical barriers to movement (i.e. land). Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent. The CTMC model then produces movement dynamics that move around barriers rather than over them. This is different than setting habitat in a patch to zero; an animal will not preferentially move towards a patch with zero habitat all else being equal, but can move over a patch with zero habitat towards another better patch. An example might be a species that lives and feeds on coastal habitats transiting through a patch of open ocean to reach a new feeding ground. In contrast, a physical barrier represents something like a peninsula that an animal must move around to transit from one side to another.

While this CTMC approach to movement simulation is to date not commonly seen in the marine modeling literature, it has numerous advantages that warrant its seeming complexity. First, when only diffusion is present, the model will generalize to the familiar dynamics of a Gaussian dispersal kernel. However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps. Second, the taxis model allow for more active habitat choices by species, allowing us to simulate both preferences of species in space and time efficiently. Lastly, this form has the advantage that the parameters are directly estimable from real data. So, if provided with say spatial abundance data and a tagging study, users can estimate the diffusion and taxis movement parameters in the same manner as @thorson2021a, and then pass their estimated parameters to `marlin` for simulation.

@@

Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry. Age is converted to length through the von Bertalanffy growth equation given parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$) assuming log-normally distributed variation *u* in the length at age with CV $\sigma_s$.

@@

We model our pelagic use case loosely on the characteristics of the Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of `r n_distinct(names(blue_fauna))` species commonly caught in the region, including both the highly abundant skipjack tuna to the heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use publicly available data on catch-per-unit-effort of each of these species from the WCPO as a proxy for baseline habitat distributions, noting that where possible, fishery-independent abundance indices would be preferable (@fig-blue).

@@

The yield impacts of the proposed MPA strategies were much more sensitive to climate impacts for the purse seine fishery. Under the *status quo* species distribution, the *Target Fishing* MPA design strategy produced increases in fisheries yields when the MPA covered 25-50% of the simulated area; that same MPA network resulted in the near complete loss of yields from the purse seine fleet when applied to the climate change scenario. In contrast, the yield outcomes for the longline fleet were relatively insensitive to the simulated range shifts (@fig-blue-results).

>>>>

Many models have been developed to support the theory and design of
spatial management strategies (see @fulton2015 for a review of existing
methods). However, they have tended to be either highly complex and
tactical models designed for use in a specific location, or to be
extremely stylized representations intended to provide theoretical
insights with less relevance for specific applications. This lack of
accessible models suited for more tactical application presents a
challenge to stakeholders charged with marine spatial planning
exercises. It also presents a barrier to the scientific community,
wherein comparisons across spatial management simulation studies are
clouded by discrepancies in underlying model structures beyond
differences in the core phenomenon in question.

To help address this challenge, we present a novel bio-economic modeling
tool called [`marlin`](https://danovando.github.io/marlin/) that allows
for efficient simulation of spatio-temporal biological and economic
dynamics. `marlin` allows users to simulate the impacts of marine
management policies across a range of species targeted by various fleets
across heterogeneous and dynamic seascapes. This model can help users
seek Pareto-optimal solutions that produce win-wins or minimize
tradeoffs across multiple management objectives, such as food production
and conservation [@lester2013].

@@

These species can be caught by fishing fleets. A fishing fleet is
defined in `marlin` by a set of fishers that have the same fishing
skill, prices, and selectivity for individual species (each of which we
denote as a *metier*). For example, both a longline and purse seine
fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the
longline fleet can be made much more likely to catch bigeye than
skipjack, and *vice versa*. Each fishing fleet distributes its fishing
effort in space according to a specified spatial allocation rule (see
*Spatial Allocation of Effort* ), for example in proportion to
profit-per-unit-effort, conditional on management policies in places
such as quotas and/or the presence of any spatial restrictions such as
MPAs.

Unconstrained by management, the total amount of effort exerted by each
fleet can follow one of two dynamics: open access or constant effort
(see *Calculating Total Effort*). Under constant effort, the total
amount of effort of each fleet is fixed over time, with the possible
exception of attrition due to MPA placement. Under open access, the
total amount of effort in a given time step is a function of the
profitability of the fleet in the previous time step, evolving until a
bionomic equilibrium of zero total profits is reached. Profitability is
a function of the volume and price of each species caught, as well as
the cost of the total amount of fishing effort per fleet and the travel
costs as a function of distance from a port (see *Fleet Dynamics*).

@@

For both the diffusion and taxis matrices, we allow for the inclusion of
physical barriers to movement (i.e. land). Pairs of patches that are
adjacent but in which one or both patches are a barrier to movement are
set as non-adjacent. The CTMC model then produces movement dynamics that
move around barriers rather than over them. This is different than
setting habitat in a patch to zero; an animal will not preferentially
move towards a patch with zero habitat all else being equal, but can
move over a patch with zero habitat towards another better patch. An
example might be a species that lives and feeds on coastal habitats
transiting through a patch of open ocean to reach a new feeding ground.
In contrast, a physical barrier represents something like a peninsula
that an animal must move around to transit from one side to another.

While this CTMC approach to movement simulation is to date not commonly
seen in the marine modeling literature, it has numerous advantages that
warrant its seeming complexity. First, when only diffusion is present,
the model will generalize to the familiar dynamics of a Gaussian
dispersal kernel. However, the time component *t* in @eq-movement allows
the movement model to be run on arbitrarily fine time steps. Second, the
taxis model allow for more active habitat choices by species, allowing
us to simulate preferences of species in space and time efficiently.
Lastly, this form has the advantage that the parameters are directly
estimable from real data. So, if provided with for example spatial
abundance data and a tagging study, users can estimate the diffusion and
taxis movement parameters in the same manner as @thorson2021a, and then
pass their estimated parameters to `marlin` for simulation.

@@

Spawning stock biomass *ssb* is calculated by converting age to mean
length, calculating weight at age, maturity at age, and then calculating
spawning stock biomass as the sum of spawning potential at age in a
given time step, taking into account the potential for hyperallometry in
the manner of @marshall2021. Age is converted to length through the von
Bertalanffy growth equation given parameters asymptotic length
($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$)
assuming log-normally distributed variation *u* in the length at age
with CV $\sigma_s$.

@@

We model our pelagic use case loosely on the characteristics of the
Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
trajectories of `r n_distinct(names(blue_fauna))` species commonly
caught in the region, including both the highly abundant skipjack tuna
and the heavily depleted oceanic whitetip shark (*Carcharhinus
longimanus*) (@fig-blue-ssb). We use publicly available data on
catch-per-unit-effort of each of these species from the WCPO as a proxy
for baseline habitat distributions, noting that where possible,
fishery-independent abundance indices would be preferable (@fig-blue).

@@

The yield impacts of the proposed MPA strategies were much more
sensitive to climate impacts for the purse seine fishery than for the longline fishery. Under the
*status quo* species distribution, the *Target Fishing* MPA design
strategy produced increases in fisheries yields when the MPA covered
25-50% of the simulated area; that same MPA network resulted in the near
complete loss of yields from the purse seine fleet when applied to the
climate change scenario. In contrast, the yield outcomes for the
longline fleet were relatively insensitive to the simulated range shifts
(@fig-blue-results).
